---
title: "Case Study 2"
author: "Daniel Chang"
date: "2022-11-19"
output: html_document
---
## Import Libraries 
Importing libraries to use for project.
```{r, echo=FALSE}
### Needed Packages
library(dplyr)
library(tidyr)
library(ggplot2)
library(class)
library(caret)
library(e1071)
library(stringr)
library(naniar)
library(rmarkdown)
library(readxl)
library(imbalance)
library(fastDummies)
library(psych)
library(GGally)
```

## Read in full employee dataset
```{r}
### Data Imports
employeeData = read.csv("CaseStudy2-data.csv", sep = ",")
employee_original = employeeData # make copy of original dataset
head(employeeData)
```

## Cleaning Dataset
In this section , I will be checking for nulls and encoding columns with characters variables to numeric variables. 
```{r}
### Check for Missing Values
sum(is.na(employeeData))

# Make overtime & Attrition column binary 
employeeData$cleanOverTime = ifelse(employeeData$OverTime=="Yes",1,0)
employeeData$cleanAttrition = ifelse(employeeData$Attrition=="Yes",1,0)
```

## Attrition Exploratory Data Analysis(EDA)
Some variables that should be included are Monthly Income, Overtime, Job Level and Job Role.
```{r}
#### Exploratory Data Analysis
#### Attrition EDA
employeeData %>% ggplot(aes(x=Attrition,fill=Attrition)) + 
  geom_bar()+
  ggtitle("Attrition Count") +
  xlab("Attrition")+ylab("Count")

### Percentage Compares for Job Role
ggplot(employeeData, aes(x = JobRole, fill = Attrition)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent)+ 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  ggtitle("Attrition Rate By Job Role")


### Attrition Vs. Job Satisfaction
employeeData %>% 
  ggplot(aes(x=JobSatisfaction,fill=Attrition))+
  geom_bar()+
  ggtitle("Attrition Vs. Job Satisfaction") 

### Attrition Vs. MonthlyIncome
employeeData %>% ggplot(aes(x=MonthlyIncome,fill=Attrition))+
  geom_histogram()+
  ggtitle("Attrition Vs. MonthlyIncome") 

### Attrition Vs. DistanceFromHome
employeeData %>% 
  ggplot(aes(x=DistanceFromHome,fill=Attrition))+
  geom_histogram()+
  ggtitle("Attrition Vs. DistanceFromHome") 

### Attrition Vs. MonthlyRate
employeeData %>% ggplot(aes(x=MonthlyRate,fill=Attrition))+geom_histogram()+ggtitle("Attrition Vs. MonthlyRate") 

### Attrition Vs. OverTime
employeeData %>% 
  ggplot(aes(x=OverTime,fill=Attrition))+
  geom_bar(position="fill")+ggtitle("Attrition Vs. Overtime")+
  scale_y_continuous(labels = scales::percent)

### Attrition Vs. years Since last Promotion
employeeData %>% 
  ggplot(aes(x=YearsSinceLastPromotion,fill=Attrition))+
  geom_bar(position="fill")+ggtitle("Attrition Vs. Years Since Last Promotion") +  
  scale_y_continuous(labels = scales::percent)

### Attrition Vs. Salary Hike - NO
employeeData %>% 
  ggplot(aes(x=PercentSalaryHike,fill=Attrition))+
  geom_histogram(position="fill")+
  ggtitle("Attrition Vs. Percent Salary Hike") 

### Attrition Vs. Age
employeeData %>% 
  ggplot(aes(x=Age,fill=Attrition))+
  geom_histogram()+
  ggtitle("Attrition Vs. Age")

### Attrition vs Marital Status
employeeData %>% 
  ggplot(aes(x=MaritalStatus,fill=Attrition))+
  geom_bar(position="fill")+
  ggtitle("Attrition Vs. Marital Status")

### Attrition Vs. PercentSalaryHike
employeeData %>% 
  ggplot(aes(x=PercentSalaryHike,fill=Attrition))+
  geom_histogram()+ggtitle("Attrition Vs. PercentSalaryHike")+
  scale_y_continuous(labels = scales::percent)

### Attrition Vs. WorkLifeBalance
employeeData %>% 
  ggplot(aes(x=WorkLifeBalance,fill=Attrition))+
  geom_histogram()+ggtitle("Attrition Vs. WorkLifeBalance")+
  scale_y_continuous(labels = scales::percent)

### Attrition Vs. StockOptionLevel
employeeData %>% 
  ggplot(aes(x=StockOptionLevel,fill=Attrition))+
  geom_histogram()+ggtitle("Attrition Vs. StockOptionLevel")+
  scale_y_continuous(labels = scales::percent)

### Attrition Vs. TrainingTimesLastYear
employeeData %>% 
  ggplot(aes(x=TrainingTimesLastYear,fill=Attrition))+
  geom_histogram()+
  ggtitle("Attrition Vs. TrainingTimesLastYear")+
  scale_y_continuous(labels = scales::percent)

### Attrition Vs. PerformanceRating
employeeData %>% 
  ggplot(aes(x=PerformanceRating,fill=Attrition))+
  geom_histogram()+ggtitle("Attrition Vs. PerformanceRating")+
  scale_y_continuous(labels = scales::percent)

### Attrition Vs. RelationshipSatisfaction
employeeData %>% 
  ggplot(aes(x=RelationshipSatisfaction,fill=Attrition))+
  geom_histogram()+ggtitle("Attrition Vs. RelationshipSatisfaction")+
  scale_y_continuous(labels = scales::percent)

### Attrition Vs. Joblevel
employeeData %>% 
  ggplot(aes(x=JobLevel,fill=Attrition))+
  geom_bar()+ggtitle("Attrition Vs. Joblevel")+
  scale_y_continuous(labels = scales::percent)

### Attrition Vs. BusinessTravel
employeeData %>% 
  ggplot(aes(x=BusinessTravel,fill=Attrition))+
  geom_bar()+ggtitle("Attrition Vs. BusinessTravel")+
  scale_y_continuous(labels = scales::percent)

#### Monthly Income EDA
### Compare Incomes
employeeData %>% group_by(Attrition) %>% summarise(compareincomes=mean(MonthlyIncome))

### Job Role vs. Monthly Salary
employeeData %>% 
  ggplot(aes(x=JobRole,y = MonthlyIncome, fill = JobRole))+
  geom_boxplot()+ggtitle("Monthly Income vs. Job Role")+
  theme(axis.text.x = element_text(angle = 60, vjust = 1, hjust=1)) 

ggpairs(employeeData[,c(3,30, 33:36)], aes(color = Attrition))

employeeData %>% 
  ggplot(aes(x=TotalWorkingYears,y = MonthlyIncome))+
  geom_point()+ggtitle("Monthly Income vs. TotalWorkingYears")+
  theme(axis.text.x = element_text(angle = 60, vjust = 1, hjust=1)) 

employeeData %>% 
  ggplot(aes(x=BusinessTravel,y = MonthlyIncome))+
  geom_boxplot()+ggtitle("Monthly Income vs. BusinessTravel")+
  theme(axis.text.x = element_text(angle = 60, vjust = 1, hjust=1))
```

## KNN Model
```{r}

### KNN Model

## Classify
employeeData2 = employeeData
employeeData2$Attrition = as.factor(employeeData2$Attrition)
employeeData2$Attrition = relevel(employeeData2$Attrition, ref = "Yes")
# create dataset for KNN model
model = employeeData2[,-c(1,4,5,6,8,9,10,11,12,13,14,17,19,21,23,24,28,38)]

# oversample to make up for imbalance in dataset
model = oversample(model,classAttr = "Attrition",method = "ADASYN")

## ML
set.seed(100)
iterations = 200
numks = 20
splitPerc = .70
masterAcc = matrix(nrow = iterations, ncol = numks)
for(j in 1:iterations)
{
  trainIndices = sample(1:dim(model)[1],round(splitPerc * dim(model)[1]))
  train = model[trainIndices,]
  test = model[-trainIndices,]
  for(i in 1:numks)
  {
    classifications = knn(train[,-c(2)],test[,-c(2)],train$Attrition, prob = TRUE, k = i)
    table(classifications,test$Attrition)
    CM = confusionMatrix(table(classifications,test$Attrition))
    masterAcc[j,i] = CM$overall[1]
  }
  
}
MeanAcc = colMeans(masterAcc)
plot(seq(1,numks,1),MeanAcc, type = "l")
which.max(MeanAcc)
max(MeanAcc)
```

## Confusion Matrix for KNN model
```{r}
classifications = knn(train[,-c(2)],test[,-c(2)],train$Attrition, prob = TRUE, k = 3)
table(classifications,test$Attrition)
CM = confusionMatrix(table(classifications,test$Attrition))
CM
```

## Naive-Bayes Model
```{r}
set.seed(25)
naive_data=employeeData

naive_data$Attrition = as.factor(naive_data$Attrition)
naive_data$Attrition = relevel(naive_data$Attrition, ref ="Yes")

model2 = naive_data[,-c(1,2,4,5,7,9,10,11,13,
                        14,21:23,25,28,37,38)]

#Convert all character variables to factors
model2 = model2 %>%
  mutate_if(sapply(model2, is.character), as.factor)
#Convert all character variables to integer 
model2 = model2 %>%
  mutate_if(sapply(model2, is.factor), as.integer)

#oversample
model2 = oversample(model2,classAttr = "Attrition",method = "ADASYN")

#revert interger back to factor
model2 = model2 %>%
  mutate_if(sapply(model2, is.integer), as.factor)

#change Attrition back to character and factor
model2$Attrition <- (ifelse(model2$Attrition==1,"No","Yes"))
model2$Attrition <- as.factor(model2$Attrition)

model2$Attrition = as.factor(model2$Attrition)
trainIndices = sample(1:dim(model2)[1],round(.70 * dim(model2)[1]))
train = model2[trainIndices,]
test = model2[-trainIndices,]

classifier1 = naiveBayes(Attrition~., data =model2)

pred = predict(classifier1,newdata=test)
CM = confusionMatrix(table(test$Attrition,pred))

CM
```


## Read in NoAttrition dataset
```{r}
employee_noattri = read.csv("CaseStudy2CompSet No Attrition (1).csv")
employee_noattri
```

## Predict Attrition
```{r}
employee_noattri$cleanOverTime = ifelse(employee_noattri$OverTime=="Yes",1,0)

attrition_pred = predict(classifier1, employee_noattri)
attrition_pred
```

## Add attrition prediction to dataset
```{r}
employee_noattri$AttritionPred = attrition_pred
employee_noattri
```
## Write csv file for attrition prediction
```{r}
filtered_noattrition = employee_noattri %>% select(ID, AttritionPred) %>% arrange(ID)
filtered_noattrition

write.csv(filtered_noattrition, "/Users/dannychang/Desktop/Case Study 2/Case2PredictionsChang Attrition.csv")
```

## Read in NoSalary dataset
```{r}
employee_nosalary = read_excel("CaseStudy2CompSet No Salary (2).xlsx")
```


## Finding Variables for Best Linear Regression Model
We will create a linear regression model to test all of the variables to decide on which one will be selected for our final linear regression model.To incoporate all the variables, I will be creating dummy columns to make various columns binary, I can use them in this model. 
```{r}
employeeData3 = employeeData2
employeeData3 = dummy_cols(employeeData3, 
                           select_columns = c("BusinessTravel","Department",
                                              "EducationField","Gender",
                                              "JobRole", "MaritalStatus" ))

employeeData3$JobRole_Others = ifelse(employeeData3$JobRole_Manager == 1|employeeData3$`JobRole_Research Director` == 1|
                                        employeeData3$`JobRole_Sales Executive` == 1, 0, 1)
employeeData3$BusinessTravel_Others = ifelse(employeeData3$`BusinessTravel_Non-Travel` == 1, 0, 1)

employeeData3[,c(39:66)] = employeeData3[,c(39:66)] %>%
  mutate_if(sapply(employeeData3[,c(39:66)], is.numeric), as.factor)

lm_salarydf = employeeData3[,c(2,5,7,8,12,14,15,16,18, 20,
                              21,25,26,27,29,30, 31,32,33,34,
                              35,36,37, 39:64)]

lmsalary_model = lm(MonthlyIncome~.,
                    data = lm_salarydf)
summary(lmsalary_model)
```

## Creating Linear Regression Model
Looking at the Linear Regression model that we created above, we will want to select all of the columns where the p-value yields a significant result. We will take those values and add it to our final linear regression model to predict salary.
```{r}
lm_salarydf = employeeData3[,c(16, 20,21, 26, 30, 35,39:44, 53:61)] # 39-44 is Business Travel, 53-61 is Jobe Role

lmsalary_model2 = lm(MonthlyIncome~., 
                    data = lm_salarydf)
summary(lmsalary_model2)

RSS = c(crossprod(lmsalary_model2$residuals))
MSE = RSS / length(lmsalary_model2$residuals)
RMSE = sqrt(MSE)
sig2 = RSS / lmsalary_model2$df.residual

RMSE
```


## Predicting Salary
```{r}
# scale no salary dataset 
#employee_nosalary$MonthlyRate=scale(employee_nosalary$MonthlyRate)
#employee_nosalary$JobSatisfaction=scale(employee_nosalary$JobSatisfaction)
#employee_nosalary$Age=scale(employee_nosalary$Age)
#employee_nosalary$EnvironmentSatisfaction=scale(employee_nosalary$EnvironmentSatisfaction)
#employee_nosalary$DailyRate=scale(employee_nosalary$DailyRate)
#employee_nosalary$JobLevel=scale(employee_nosalary$JobLevel)
#employee_nosalary$PercentSalaryHike=scale(employee_nosalary$PercentSalaryHike)

# create dummy variables
employee_nosalary= dummy_cols(employee_nosalary, 
                           select_columns = c("BusinessTravel","Department",
                                              "EducationField","Gender",
                                              "JobRole", "MaritalStatus" ))

employee_nosalary$cleanOverTime = ifelse(employee_nosalary$OverTime=="Yes",1,0)
employee_nosalary$JobRole_Others = ifelse(employee_nosalary$JobRole_Manager == 1|
                                            employee_nosalary$`JobRole_Research Director` ==1|
                                            employee_nosalary$`JobRole_Sales Executive` == 1, 0, 1)
employee_nosalary$BusinessTravel_Others = ifelse(employee_nosalary$`BusinessTravel_Non-Travel` == 1, 0, 1)

# change dummy variables to factors
employee_nosalary[,c(36:63)] = employee_nosalary[,c(36:63)] %>%
  mutate_if(sapply(employee_nosalary[,c(36:63)], is.numeric), as.factor)

# predict salary
salary_pred = predict(lmsalary_model2, employee_nosalary)
#sd = sd(employee_original$MonthlyIncome)
#salary_mean = mean(employee_original$MonthlyIncome)
#unscaled_pred = (salary_pred * sd) + salary_mean
#unscaled_pred 
```

## Adding Column to No Salary Dataset and Exporting as .csv File
```{r}
#salary_pred = unlist(salary_pred)

employee_nosalary$SalaryPred =salary_pred 
head(employee_nosalary)

filtered_nosalary = employee_nosalary %>% select(ID, SalaryPred) %>% arrange(ID)
filtered_nosalary
```
## Create csv file with salary predictions
```{r}
write.csv(filtered_nosalary, "/Users/dannychang/Desktop/Case Study 2/Case2PredictionsChang Salary.csv")
```

