---
title: "Case Study 2"
author: "Daniel Chang"
date: "2022-11-19"
output: html_document
---
# Employee Attrition and Salary Analysis and Prediction Models
## Import Libraries 
Importing libraries to use for project.
```{r, echo=FALSE}
### Needed Packages
library(dplyr)
library(tidyr)
library(ggplot2)
library(class)
library(caret)
library(e1071)
library(stringr)
library(naniar)
library(rmarkdown)
library(readxl)
library(imbalance)
library(fastDummies)
library(psych)
library(GGally)
```

## Read in full employee dataset
```{r}
### Data Imports
employeeData = read.csv("CaseStudy2-data.csv", sep = ",")
employee_original = employeeData # make copy of original dataset
head(employeeData)
```

## Cleaning Dataset
In this section , I will be checking for nulls and encoding columns with characters variables to numeric variables. 
There is nothing to clean here as there are no nulls in the dataset and the data is high in integrity. 
```{r}
### Check for Missing Values
sum(is.na(employeeData))

# Make overtime & Attrition column binary 
employeeData$cleanOverTime = ifelse(employeeData$OverTime=="Yes",1,0)
employeeData$cleanAttrition = ifelse(employeeData$Attrition=="Yes",1,0)
```

## Attrition Exploratory Data Analysis(EDA)
First, we started of will looking at the attrition count. From our bar chart, we find that the "No" Attrition outnumbers the "Yes" by nearly 7 to 1.

Some variables that should be included are Monthly Income, Overtime, Job Level and Job Role.

Our Attrition vs. Monthly Income histogram shows that as the salary of an individual increases, they are less likely to quit. Next, we find that those who work overtime are more likely ot quit. A little over 25% of those who work overtime quite which makes sense since they are more prone to stress. Job Role seems to be a big factor in those who quite as well. Almost 50% of those who work as sales representative quit and about 25% of those who work in human resources do. 

Furthermore, from our pair plots, we find that the length some works is somewhat correlated with the attrition rate as well. The correlation hovers between 0.59 and0.78 which a strong and positive correlation, not very strong. However, it is enough to make an impact.   
```{r}
#### Exploratory Data Analysis
#### Attrition EDA
employeeData %>% ggplot(aes(x=Attrition,fill=Attrition)) + 
  geom_bar()+
  ggtitle("Attrition Count") +
  xlab("Attrition")+ylab("Count")

### Percentage Compares for Job Role
ggplot(employeeData, aes(x = JobRole, fill = Attrition)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent)+ 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  ggtitle("Attrition Rate By Job Role")


### Attrition Vs. Job Satisfaction
employeeData %>% 
  ggplot(aes(x=JobSatisfaction,fill=Attrition))+
  geom_bar()+
  ggtitle("Attrition Vs. Job Satisfaction") 

### Attrition Vs. MonthlyIncome
employeeData %>% ggplot(aes(x=MonthlyIncome,fill=Attrition))+
  geom_histogram()+
  ggtitle("Attrition Vs. MonthlyIncome") 

### Attrition Vs. DistanceFromHome
employeeData %>% 
  ggplot(aes(x=DistanceFromHome,fill=Attrition))+
  geom_histogram()+
  ggtitle("Attrition Vs. DistanceFromHome") 

### Attrition Vs. MonthlyRate
employeeData %>% ggplot(aes(x=MonthlyRate,fill=Attrition))+geom_histogram()+ggtitle("Attrition Vs. MonthlyRate") 

### Attrition Vs. OverTime
employeeData %>% 
  ggplot(aes(x=OverTime,fill=Attrition))+
  geom_bar(position="fill")+ggtitle("Attrition Vs. Overtime")+
  scale_y_continuous(labels = scales::percent)

### Attrition Vs. years Since last Promotion
employeeData %>% 
  ggplot(aes(x=YearsSinceLastPromotion,fill=Attrition))+
  geom_bar(position="fill")+ggtitle("Attrition Vs. Years Since Last Promotion") +  
  scale_y_continuous(labels = scales::percent)

### Attrition Vs. Salary Hike - NO
employeeData %>% 
  ggplot(aes(x=PercentSalaryHike,fill=Attrition))+
  geom_histogram(position="fill")+
  ggtitle("Attrition Vs. Percent Salary Hike") 

### Attrition Vs. Age
employeeData %>% 
  ggplot(aes(x=Age,fill=Attrition))+
  geom_histogram()+
  ggtitle("Attrition Vs. Age")

### Attrition vs Marital Status
employeeData %>% 
  ggplot(aes(x=MaritalStatus,fill=Attrition))+
  geom_bar(position="fill")+
  ggtitle("Attrition Vs. Marital Status")

### Attrition Vs. PercentSalaryHike
employeeData %>% 
  ggplot(aes(x=PercentSalaryHike,fill=Attrition))+
  geom_histogram()+ggtitle("Attrition Vs. PercentSalaryHike")+
  scale_y_continuous(labels = scales::percent)

### Attrition Vs. WorkLifeBalance
employeeData %>% 
  ggplot(aes(x=WorkLifeBalance,fill=Attrition))+
  geom_histogram()+ggtitle("Attrition Vs. WorkLifeBalance")+
  scale_y_continuous(labels = scales::percent)

### Attrition Vs. StockOptionLevel
employeeData %>% 
  ggplot(aes(x=StockOptionLevel,fill=Attrition))+
  geom_histogram()+ggtitle("Attrition Vs. StockOptionLevel")+
  scale_y_continuous(labels = scales::percent)

### Attrition Vs. TrainingTimesLastYear
employeeData %>% 
  ggplot(aes(x=TrainingTimesLastYear,fill=Attrition))+
  geom_histogram()+
  ggtitle("Attrition Vs. TrainingTimesLastYear")+
  scale_y_continuous(labels = scales::percent)

### Attrition Vs. PerformanceRating
employeeData %>% 
  ggplot(aes(x=PerformanceRating,fill=Attrition))+
  geom_histogram()+ggtitle("Attrition Vs. PerformanceRating")+
  scale_y_continuous(labels = scales::percent)

### Attrition Vs. RelationshipSatisfaction
employeeData %>% 
  ggplot(aes(x=RelationshipSatisfaction,fill=Attrition))+
  geom_histogram()+ggtitle("Attrition Vs. RelationshipSatisfaction")+
  scale_y_continuous(labels = scales::percent)

### Attrition Vs. Joblevel
employeeData %>% 
  ggplot(aes(x=JobLevel,fill=Attrition))+
  geom_bar()+ggtitle("Attrition Vs. Joblevel")+
  scale_y_continuous(labels = scales::percent)

### Attrition Vs. BusinessTravel
employeeData %>% 
  ggplot(aes(x=BusinessTravel,fill=Attrition))+
  geom_bar()+ggtitle("Attrition Vs. BusinessTravel")+
  scale_y_continuous(labels = scales::percent)

#### Monthly Income EDA
### Compare Incomes
employeeData %>% group_by(Attrition) %>% summarise(compareincomes=mean(MonthlyIncome))

### Job Role vs. Monthly Salary
employeeData %>% 
  ggplot(aes(x=JobRole,y = MonthlyIncome, fill = JobRole))+
  geom_boxplot()+ggtitle("Monthly Income vs. Job Role")+
  theme(axis.text.x = element_text(angle = 60, vjust = 1, hjust=1)) 

ggpairs(employeeData[,c(3,30, 33:36)], aes(color = Attrition))

employeeData %>% 
  ggplot(aes(x=TotalWorkingYears,y = MonthlyIncome))+
  geom_point()+ggtitle("Monthly Income vs. TotalWorkingYears")+
  theme(axis.text.x = element_text(angle = 60, vjust = 1, hjust=1)) 

employeeData %>% 
  ggplot(aes(x=BusinessTravel,y = MonthlyIncome))+
  geom_boxplot()+ggtitle("Monthly Income vs. BusinessTravel")+
  theme(axis.text.x = element_text(angle = 60, vjust = 1, hjust=1))
```

## KNN Model
```{r}
### KNN Model

## Classify
employeeData2 = employeeData
employeeData2$Attrition = as.factor(employeeData2$Attrition)

# create dataset for KNN model
model = employeeData2[,-c(1,4,5,6,8,9,10,11,12,13,14,17,19,21,23,24,28,38)]

# oversample to make up for imbalance in dataset
model = oversample(model,classAttr = "Attrition",method = "ADASYN")

## ML
set.seed(124) # Changed seed multiple times to see how high it can go
iterations = 200
numks = 20
splitPerc = .70
masterAcc = matrix(nrow = iterations, ncol = numks)
for(j in 1:iterations)
{
  trainIndices = sample(1:dim(model)[1],round(splitPerc * dim(model)[1]))
  train = model[trainIndices,]
  test = model[-trainIndices,]
  for(i in 1:numks)
  {
    classifications = knn(train[,-c(2)],test[,-c(2)],train$Attrition, prob = TRUE, k = i)
    table(classifications,test$Attrition)
    CM = confusionMatrix(table(classifications,test$Attrition))
    masterAcc[j,i] = CM$overall[1]
  }
  
}
MeanAcc = colMeans(masterAcc)
plot(seq(1,numks,1),MeanAcc, type = "l")
which.max(MeanAcc)
max(MeanAcc)
```

## Confusion Matrix for KNN model
As we build and analyze our KNN model, we find that KNN is not the best model to use. Our accuracy is abysmal. Our highest sensitivity is only slightly better at .6471 while our specificity is worse at 0.7719. As a result, we will move on to a different model.   
```{r}
classifications = knn(train[,-c(2)],test[,-c(2)],train$Attrition, prob = TRUE, k = 3)
table(classifications,test$Attrition)
CM = confusionMatrix(table(classifications,test$Attrition))
CM
```

## Naive-Bayes Model
Now, want to use Naive-bayes to build our model. This model yielded us much better results with an accuracy of 0.8884, a Specificity of 0.9019 and sensitivity 0.8750. As a result, we find this model to be quite accurate. Therefore, the Naive-bayes model that we've built to predict attriton on the test dataset.   
```{r}
set.seed(25)
naive_data=employeeData

naive_data$Attrition = as.factor(naive_data$Attrition)

model2 = naive_data[,-c(1,2,4,5,7,9,10,11,13,
                        14,21:23,25,28,37,38)]

## get data ready for oversampling
#Convert all character variables to factors
model2 = model2 %>%
  mutate_if(sapply(model2, is.character), as.factor)
#Convert all character variables to integer 
model2 = model2 %>%
  mutate_if(sapply(model2, is.factor), as.integer)

#oversample
model2 = oversample(model2,classAttr = "Attrition",method = "ADASYN")

#revert interger back to factor
model2 = model2 %>%
  mutate_if(sapply(model2, is.integer), as.factor)

#change Attrition back to character and factor
model2$Attrition <- (ifelse(model2$Attrition==1,"No","Yes"))
model2$Attrition <- as.factor(model2$Attrition)

model2$Attrition = as.factor(model2$Attrition)
trainIndices = sample(1:dim(model2)[1],round(.70 * dim(model2)[1]))
train = model2[trainIndices,]
test = model2[-trainIndices,]

classifier1 = naiveBayes(Attrition~., data =model2)

pred = predict(classifier1,newdata=test)
CM = confusionMatrix(table(test$Attrition,pred))

CM
```


## Read in NoAttrition dataset
```{r}
employee_noattri = read.csv("CaseStudy2CompSet No Attrition (1).csv")
employee_noattri
```

## Predict Attrition
In this code block, we will use our Naive-Bayes model to predict attrition and store it in a variable. 
```{r}
# Convert overtime to binary 
employee_noattri$cleanOverTime = ifelse(employee_noattri$OverTime=="Yes",1,0)

employee_noattri = employee_noattri %>%
  mutate_if(sapply(employee_noattri, is.character), as.factor)
#Convert all character variables to integer 
employee_noattri = employee_noattri %>%
  mutate_if(sapply(employee_noattri, is.factor), as.integer)
employee_noattri = employee_noattri %>%
  mutate_if(sapply(employee_noattri, is.integer), as.factor)


attrition_pred = predict(classifier1, employee_noattri)
attrition_pred
```

## Add attrition prediction to dataset
```{r}
employee_noattri$AttritionPred = attrition_pred
employee_noattri
```

## Write csv file for attrition prediction
```{r}
filtered_noattrition = employee_noattri %>% 
  select(ID, AttritionPred) %>% arrange(ID)
filtered_noattrition

write.csv(filtered_noattrition, "Case2PredictionsChang Attrition.csv")
```

## Read in NoSalary dataset
```{r}
employee_nosalary = read_excel("CaseStudy2CompSet No Salary (2).xlsx")
```


## Finding Variables for Best Linear Regression Model
We will create a linear regression model to test all of the variables to decide on which one will be selected for our final linear regression model.To incoporate all the variables, I will be creating dummy columns to make various columns binary, I can use them in this model. 
```{r}
employeeData3 = employeeData2
employeeData3 = dummy_cols(employeeData3, 
                           select_columns = c("BusinessTravel","Department",
                                              "EducationField","Gender",
                                              "JobRole", "MaritalStatus" ))

employeeData3$JobRole_Others = ifelse(employeeData3$JobRole_Manager == 1|employeeData3$`JobRole_Research Director` == 1|
                                        employeeData3$`JobRole_Sales Executive` == 1, 0, 1)
employeeData3$BusinessTravel_Others = ifelse(employeeData3$`BusinessTravel_Non-Travel` == 1, 0, 1)

employeeData3[,c(39:66)] = employeeData3[,c(39:66)] %>%
  mutate_if(sapply(employeeData3[,c(39:66)], is.numeric), as.factor)

lm_salarydf = employeeData3[,c(2,5,7,8,12,14,15,16,18, 20,
                              21,25,26,27,29,30, 31,32,33,34,
                              35,36,37, 39:64)]

lmsalary_model = lm(MonthlyIncome~.,
                    data = lm_salarydf)
summary(lmsalary_model)
```

## Creating Linear Regression Model
Looking at the Linear Regression model that we created above, we will want to select all of the columns where the p-value yields a significant result. We will take those values and add it to our final linear regression model to predict salary. Please note that column 39-44 indicates all the Business travel levels while 53-61 refers to all the Job Role.
```{r}
lm_salarydf = employeeData3[,c(16, 20,21, 26, 30, 35,39:44, 53:61)] # 39-44 is Business Travel, 53-61 is Jobe Role

lmsalary_model2 = lm(MonthlyIncome~., 
                    data = lm_salarydf)
summary(lmsalary_model2)

RSS = c(crossprod(lmsalary_model2$residuals))
MSE = RSS / length(lmsalary_model2$residuals)
RMSE = sqrt(MSE)
sig2 = RSS / lmsalary_model2$df.residual

RMSE
```


## Predicting Salary
```{r}
# create dummy variables
employee_nosalary= dummy_cols(employee_nosalary, 
                           select_columns = c("BusinessTravel","Department",
                                              "EducationField","Gender",
                                              "JobRole", "MaritalStatus" ))

employee_nosalary$cleanOverTime = ifelse(employee_nosalary$OverTime=="Yes",1,0)
employee_nosalary$JobRole_Others = ifelse(employee_nosalary$JobRole_Manager == 1|
                                            employee_nosalary$`JobRole_Research Director` ==1|
                                            employee_nosalary$`JobRole_Sales Executive` == 1, 0, 1)
employee_nosalary$BusinessTravel_Others = ifelse(employee_nosalary$`BusinessTravel_Non-Travel` == 1, 0, 1)

# change dummy variables to factors
employee_nosalary[,c(36:63)] = employee_nosalary[,c(36:63)] %>%
  mutate_if(sapply(employee_nosalary[,c(36:63)], is.numeric), as.factor)

# predict salary
salary_pred = predict(lmsalary_model2, employee_nosalary)
```

## Adding Column to No Salary Dataset and Exporting as .csv File
```{r}
#salary_pred = unlist(salary_pred)

employee_nosalary$SalaryPred =salary_pred 
head(employee_nosalary)

filtered_nosalary = employee_nosalary %>% select(ID, SalaryPred) %>% arrange(ID)
filtered_nosalary
```
## Create csv file with salary predictions
```{r}
write.csv(filtered_nosalary, "Case2PredictionsChang Salary.csv")
```

